---
title: navigating SPSS data with purrr and stringr - part i
author: renata gerecke
date: '2019-01-20'
categories:
  - haven
  - purrr
  - rstats
  - spss
  - stringr
tags:
  - haven
  - purrr
  - rstats
  - spss
  - stringr
slug: navigating-spss-data-with-purrr-and-stringr
weight: 11
---

[**Download data for this blog**](https://github.com/rvg02010/blog/tree/master/static/data/2019-01-20_sample-data.sav)

There is a lot not to like about SPSS: it's expensive, it's proprietary, it's not very flexible. But I have to give it credit for one thing: maintaining separate variable names and labels. 

In SPSS data files--which use the `.sav` extension--each variable has a name like an R variable. These names follow many of the same conventions of R variable names, for example, they can't contain spaces or start with numbers. But each SPSS data file also has a built-in data dictionary, which includes:

- A label (which can include spaces, punctuation)
- Value labels
- Variable type
- Other metadata

The `haven` package allows you to easily load in SPSS data files--including that sweet, sweet variable metadata. 

```{r read data}
suppressPackageStartupMessages(library(tidyverse))
library(haven)

df <- read_sav("../../static/data/2019-01-20_sample-data.sav")

glimpse(df)
```

In R, having variables labeled `q1` through `q10` can be quite nice--for example, it allows you to take advantage of `dplyr::num_range()` or `dplyr::starts_with()` when trying to select particular variables. But they don't really give you a sense of what the variable represents. 

By reading in the labelled SPSS data with the `haven` package, however, we have preserved the variable labels! For a long time, I had no idea how to access this information--it took reading the first chapter of [Hadley Wickham's Advanced R](http://adv-r.had.co.nz/) (in particular, [this section on attributes](http://adv-r.had.co.nz/Data-structures.html#attributes)) for me to figure it out. 

Here, he explains that we can use the function `attributes()` to call an object's attributes--in this case, the metadata for a given SPSS variable:

```{r metadata for q1}
attributes(df$q1)
```


If we want to call a specific attribute--for example, just the label of a given variable--we can use the function `attr()` like so:

```{r label for q1}
attr(df$q1, "label")
```

And if we want to read the labels for all of our variables? We can quickly iterate through our data frame using `purrr::map_chr()` [which handily returns a character vector instead of a list]:

```{r labels for df}
map_chr(df, ~attr(., "label"))
```

Tune in next time to learn how to use `stringr` to use SPSS variable labels to transform our data!
