---
title: navigating SPSS variables with purrr and stringr - part ii
author: renata gerecke
date: '2019-02-03'
categories:
  - haven
  - purrr
  - rstats
  - spss
  - stringr
tags:
  - haven
  - purrr
  - rstats
  - spss
  - stringr
slug: navigating-spss-variables-with-purrr-and-stringr-part-ii
---

[**Download data for this blog**](https://github.com/rvg02010/blog/tree/master/static/data/2019-01-20_sample-data.sav)

## Recap

[Last post](https://renatagerecke.netlify.com/post/navigating-spss-data-with-purrr-and-stringr/), I wrote about how we can access SPSS variable metadata using the `attributes()` and `attr()` functions in R; in combination with `purrr:map_chr()` we can easily view variable labels, levels, and more. 

Though the sample data I've created only has 10 variables, often I find myself working with hundreds, and little idea how to find the precise variable names I'm looking for. I've found `stringr:str_detect()` (in conjuction with `dplyr::select_if()`) to be particularly powerful in this scenario. 

The data we have asks respondents various combinations of two things: to what degree do you [share/read] on [given social network]. Let's say I want to find the questions which ask about Twitter. I can use the following:

```{r search for twitter}
suppressPackageStartupMessages(library(tidyverse))
library(haven)

df <- read_sav("../../static/data/2019-01-20_sample-data.sav")

df %>%
  select_if(~attr(., "label") %>% str_detect("Twitter"))

```


We can now see that `q3` and `q8` are the questions which ask about Twitter behavior. We could also use this message to select the columns relating to where respondents like to "post":

```{r}
df %>%
  select_if(~attr(., "label") %>% str_detect("post"))
```

If you just want to analyze a subset of these variables at a time, this might be sufficient for your needs. However, sometimes you might want to use information in these labels as another variable. As far as I know, there isn't a straightforward way to do this so here's my approach:

1. Create a separate table with variable name/label pairs using `purrr::enframe()`
2. `tidyr::gather()` the variables you're interested in working with
3. Join the two data sets
4. Mutate the original labels as necessary for analysis

To illustrate, I'm going to edit our data so we can chart average interest in each social media platform, grouped by whether it's interest in posting or reading. 

```{r}
data_dictionary <- df %>%
  map_chr(~attr(., "label")) %>%
  enframe(name = "variable", value = "label")

data_dictionary

df_gathered <- df %>%
  gather("variable", "rating", -respid) %>%
  left_join(data_dictionary) %>%
  separate(label, str_glue("word{1:7}")) %>%
  mutate(platform = factor(word1),
         activity = factor(word7)) %>%
  select(-num_range("word", 1:7))

df_gathered

df_grouped <- df_gathered %>%
  group_by(activity, platform) %>%
  summarise(rating = mean(rating))

df_grouped
```

And now we plot!

```{r}
ggplot(df_grouped) +
  aes(x = platform,
      y = rating,
      group = activity,
      color = activity) +
  geom_line(stat = "identity") +
  theme_minimal()
```

